Multi sky-view\\ 3D aerosol distribution recovery

### 2013-10-03 article.tex
Yoav modifications

The MC simulations rely on the same grid as
\cref{sec:single-scatt-model}. We use backward MC: from a detector
pixel, photons are `launched' via the camera lens; then the photons
are traced back through the atmosphere. Photons that happen to be
back-traced to the Sun are counted, per pixel. In this model, RT takes
the following steps, per pixel (\cref{fig:mcgrid}a):
\begin{figure}
  \centering
  \yoavcomment{\includegraphics{images/mcgrid.pdf}}
  \caption{\small [a] Multiple-scattering forward model by MC.
    [b] Photograph of the same sky and view point as in
    \cref{fig:simulation-results1}(a), rendered using MC.}
  \label{fig:mcgrid}
\end{figure}

@@

The quality of MC increases with the number of photons launched.
\Cref{fig:mcgrid}(b) shows a photograph of the {\em Haze Blobs}
scene, derived by the MC process, using \num{10000} photons per pixel,
from the same viewpoint as \cref{fig:simulation-results1}(a).
Figure~\ref{fig:cross_sections} plots the intensity profile in two lateral
cross sections of the viewed field, extracted from the images in
\cref{fig:simulation-results1}(d) and the corresponding MC derived photograph.
These plots demonstrate consistency in these cases, for most pixels.
\begin{figure}[t]
  \centering
  \yoavcomment{\includegraphics{images/cross_sections.pdf}}
  \caption{\small Cross-sections along the $X$-axis of photographs rendered
    by the single-scattering and MC forward models. 
    (a) cross sections of low density simulations (\cref{fig:simulation-results1}(d)
    and \cref{fig:simulation_particles}(a)). (b) cross sections of
    high density simulations (\cref{fig:simulation_particles}(d)).
  }
  \label{fig:cross_sections}
\end{figure}

\Cref{fig:scatter_order} shows the contribution by successive
scattering orders. They are derived from the MC image. Photons contributing
to an image pixel are accumulated in steps {\tt ii},{\tt v} above.
The contributing photons that stem from just a single scattering event yield \cref{fig:scatter_order}(a). Similarly, contributing photons that had experienced exactly  two or three scattering events yield respectively \cref{fig:scatter_order}(b) and
\cref{fig:scatter_order}(c). Clearly, first-order scatter is responsible
for most of the energy in the MC image. Radiance by higher order scattering
is mostly evident at the horizon. A horizontal LOS passes through a long and dense part of the atmosphere, while a vertical LOS cuts short through the dense, lower part. Hence,
towards the horizon the probability of high order scattering increases.
\begin{figure}[bt]
  \centering
  \yoavcomment{\includegraphics{images/scatter_order.pdf}}
  \caption{\small Separation of the image shown in \cref{fig:mcgrid}(b) to contribution by successive orders of scattering: (a) first , (b) second, (c) third and (d) forth scattering order.}
  \label{fig:scatter_order}
\end{figure}

@@

To demonstrate recovery, we performed simulations. We tested how results are affected by the following variations:
$\{1\}$ Multiple or only single scatter in the forward model;
$\{2\}$ Spatial distribution;
$\{3\}$ Density of viewpoints;
$\{4\}$ Aerosol characteristics;
$\{5\}$ Density scale.

In all cases, images were created as described
in \cref{sec:simul,sec:monte-carlo-simul}. These images served as input to the reconstruction algorithm. The optimization used an L-BFGS-B solver~\cite{BFGS} on a computer cluster. Each computer core was dedicated to rendering a modeled image. The algorithm was initialized
by ${\bm n}=0$.  Satisfactory convergence occurred in several hundred
iterations. Depending on the resolution it took between minutes to a
couple of hours to complete, in total. The total estimation
error can be quantified by the aerosol mass that is over and under-estimated
in all voxels, relative to the total aerosol mass in the scene. Using the $\ell_1$ norm,
the total mass relative difference is
$(\|\hat{\bm n}\|_1 - \|{\bm n}^{\rm true} \|_1) / \| {\bm n}^{\rm true} \|_1$. 
To also sense local errors, we use 
 $\epsilon=\|
\hat{\bm n} - {\bm n}^{\rm true} \|_1 / \| {\bm n}^{\rm true} \|_1$.
%  In Figs.~\ref{fig:results}(a) and (b), respectively $\epsilon=0.14$ and
%$\epsilon=0.20$.  In Figs.~\ref{fig:results}(c) and (d), both cases have
%$\epsilon=0.65$.
These results are listed in Table~\ref{tbl:mass}. \\

\noindent{\bf Forward models and distributions}: 
\Cref{fig:results} shows estimation results, corresponding to the
original distributions of \cref{fig:distributions}.
\begin{figure}[h]
  \centering
  \yoavcomment{\includegraphics{images/results}}
  \caption{\small Reconstructions of the scenes shown in
    \cref{fig:distributions}. Color represents aerosol density. The
    density units are $10^{6}~{\rm particles}/{\rm m}^3.$ In [a,b],
    single-scattering simulated images are used as input to the
    recovery.  In [c,d], MC simulated images are used as input for the
    recovery.  }
  \label{fig:results}
\end{figure}
\hl{
In these simulations the atmosphere domain was defined over a
$20\times20 \times 40$ grid. Thirty six cameras were placed
$\sim7$km apart on a $6\times6$ grid.
}
In Figs.~\ref{fig:results}(a) and (b), the single-scattering model was
used both in rendering of the raw images, and in the estimation algorithm. In
Figs.~\ref{fig:results}(c) and (d), the MC model (multiple scattering)
rendered the raw images, over which the estimation algorithm
ran. Overall, the distributions, the density order of magnitude and
values are consistent.
Errors in Figs.~\ref{fig:results}(a) and (b) stem from random noise, which had
been induced in the raw images. Errors are larger when MC is used to
render the images, since high-order scattering is not accounted for in
the algorithm of \cref{sec:inverse-problem}.
The overall mass is recovered, but for 6.5\% loss. \\

% at various grid
%resolutions.  Depending directly on the area grid finesse, $N_{\rm
%  views}$ varied from 12 to 100.
%
%
%For local errors,
% $\epsilon=\|
%\hat{\bm n} - {\bm n}^{\rm true} \|_1 / \| {\bm n}^{\rm true} \|_1$.
%  In Figs.~\ref{fig:results}(a) and (b), respectively $\epsilon=0.14$ and
%$\epsilon=0.20$.  In Figs.~\ref{fig:results}(c) and (d), both cases have
%$\epsilon=0.65$.

\noindent{\bf Density of viewpoints}: Keeping the domain and scene ({\em Haze blobs}) fixed, we
repeated the reconstruction many times, each time using a different subset of the above mentioned 36 simulated cameras. Each test randomly selected viewpoints, for various fixed values
of $N_{\rm views}< 36$. All rendered images used MC.
\Cref{fig:error_vs_cameras} plots $\epsilon$ as a function $N_{\rm views}$. 
\begin{figure}[t]
  \centering
  \yoavcomment{\includegraphics{images/error_vs_cameras.pdf}}
  \caption{\small Relative error $\epsilon$ measure as function of the number of
    cameras. Bars represent the standard deviation of $\epsilon$, over our random tests.}
  \label{fig:error_vs_cameras}
\end{figure}
Diminishing returns are obtained beyond $N_{\rm views}\approx 20$, which corresponds to $\approx 11$km separation between neighboring viewpoints.\\

\noindent{\bf Aerosol characteristics}: We used the {\em Haze blobs} scene, but varied the  
aerosol characteristics. Compared to the aerosol described in \cref{sec:simul}, we tested two variations. In one variation, the phase function was isotropic, i.e., 
$[g_{\rm R},g_{\rm G},g_{\rm B}]=[0,0,0]$. In the other variation, the aerosol is partly absorbing: its single scattering albedo per color channel is 
$[\varpi^{\rm aerosol}_{\rm R},\varpi^{\rm aerosol}_{\rm G},\varpi^{\rm aerosol}_{\rm B}]=
[0.69,0.74,0.78]$.  All rendered images used MC, and recovery used $N_{\rm views}=36$.
Sample results are shown in Figs.~\ref{fig:simulation_particles}(a)--(d) and in Table~\ref{tbl:mass}.\\

\newcolumntype{g}{>{\columncolor{Gray}}r}
\begin{table}[htbp]
  \centering
  \begin{tabular}{cccccg}
    \toprule
    Haze Scene & Forward Model& Density & Isotropic PF & Absorbing & $\epsilon$ \\
    \midrule
    Front & Single & Low & & & 0.18 \\
     & MC & Low & & & 0.65 \\
    \cmidrule(r){1-5}
    Blobs & Single & Low & & & 0.14 \\
     & MC & Low & & & 0.65 \\
     & MC & Low & $\surd$  & & 0.49 \\
     & MC & Low & & $\surd$ & 0.83 \\
     & MC & High & & & 0.98 \\
    \bottomrule
  \end{tabular}
  \caption{Local error for the various reconstruction simulations}
  \label{tbl:mass}
\end{table}

\noindent{\bf Density scale}: We used the aerosol {\em Haze blobs} distribution
and characteristics as described in Sec.~\ref{sec:simul}, but increased the
aerosol density tenfold everywhere. This significantly increases the effects
of high order scattering, as shown in \cref{fig:cross_sections}(b) and
\cref{fig:simulation_particles}(c). Consequently, $\epsilon$ increases,
when using $N_{\rm views}=36$. Still, \cref{fig:reconstructions}(c) shows
that the spatial distribution outlines are recovered similarly to the other
tests.

%\hl{Thickening the aerosols density ten folds in the} {\em Haze Blobs} \hl{scene
%resulted in a larger local error, i.e. $\epsilon=0.98$ (compared to $\epsilon=0.65$).
%This admits to the fact that the single-scattering model, on which the reconstruction
%algorithm is based, is valid when the optical depth of the medium is small.}
\begin{figure}[h]
  \centering
  \yoavcomment{\includegraphics{images/simulation_particles.pdf}}
  \caption{\small (a) and (b) show MC simulations of low and
    high density scenes, respectively. (c) shows the reconstruction of
    the high density scene.}
  \label{fig:simulation_particles}
\end{figure}
\begin{figure}[h]
  \centering
  \yoavcomment{\includegraphics{images/reconstructions.pdf}}
  \caption{\small (a) and (b) show MC simulations of low and
    high density scenes, respectively. (c) shows the reconstruction of
    the high density scene.}
  \label{fig:reconstructions}
\end{figure}

@@

We describe a way to sense the 3D atmosphere. It can help remote
sensing departing from common 1D aerosol distribution models, in favor
of detailed 3D RT analysis.  The paper describes both a novel data
acquisition system concept for this recovery task (ground-based camera
network), and a dedicated algorithm for reconstructing aerosol
distributions. Using more prior knowledge on the nature of
distributions will improve the recovery.  The estimation algorithm
uses a single scattering model. This approximation is valid when the
optical depth is small. However, as the MC simulation show, it is
important to generalize this algorithm so that it explicitly
incorporates high-order scattering in the scene, including the
ground. This will certainly be needed for dense distributions as
clouds. \Cref{fig:scatter_order}(b) \hl{suggests that incorporating
the second order of scattering might be sufficient for most situations.}

To deploy such a system on a very large scale will require a long-term
engineering effort. Camera specifications~\cite{Pust2011} should be
set to optimally recover a wide range of aerosol distributions.
\hl{The relation between the quantity and spatial distribution of
  cameras and the quality of reconstruction should also be determined.}
\fix{This requires}{These open questions require} a major theoretic
 and numerical research.

@@

>>>>

The MC simulations rely on the same grid as
\cref{sec:single-scatt-model}. We use backward MC: from a detector
pixel, photons are `launched' via the camera lens; then the photons
are traced back through the atmosphere. Photons that happen to be
back-traced to the Sun are counted, per pixel. In this model, RT takes
the following steps, per pixel (\cref{fig:mcgrid}a):
\begin{figure}
  \centering
  \yoavcomment{\includegraphics{images/mcgrid.pdf}}
  \caption{\small [a] Multiple-scattering forward model by MC.
    [b] Photograph of the same sky and view point as in
    \cref{fig:simulation-results1}[left], rendered using MC.}
  \label{fig:mcgrid}
\end{figure}

@@

The quality of MC increases with the number of photons launched.
\Cref{fig:mcgrid}(b) shows a photograph of the {\em Haze Blobs}
scene, derived by the MC process, using \num{10000} photons per pixel,
from the same viewpoint as \cref{fig:simulation-results1}[left].
Figure~\ref{fig:cross_sections} plots the intensity profile in two lateral
cross sections of the viewed field, extracted from the images in \cref{fig:simulation-results1}[left] and \cref{fig:mcgrid}(b).
These plots demonstrate consistency in these cases, for most pixels.
\begin{figure}[t]
  \centering
  \yoavcomment{\includegraphics{images/cross_sections.pdf}}
  \caption{\small Cross sections along the $X$-axis [a] and $Y$-axis [b] of
    and images rendered by either single-scattering and MC. The corresponding
    images appear in \cref{fig:simulation-results1}[left] and \cref{fig:mcgrid}(b).}
  \label{fig:cross_sections}
\end{figure}

\Cref{fig:scatter_order} shows the contribution by successive
scattering orders. They are derived from the MC image. Photons contributing to an image pixel are accumulated in steps {\tt ii},{\tt v} above.
The contributing photons that stem from just a single scattering event yield \cref{fig:scatter_order}(a). Similarly, contributing photons that had experienced exactly  two or three scattering events yield respectively \cref{fig:scatter_order}(b) and
\cref{fig:scatter_order}(c). Clearly, first-order scatter is responsible
for most of the energy in the MC image. Radiance by higher order scattering
is mostly evident at the horizon. A horizontal LOS passes through a long and dense part of the atmosphere, while a vertical LOS cuts short through the dense, lower part. Hence,
towards the horizon the probability of high order scattering increases.
\begin{figure}[bt]
  \centering
  \yoavcomment{\includegraphics{images/scatter_order.pdf}}
  \caption{\small Separation of the image shown in \cref{fig:mcgrid}(b) to contribution by successive orders of scattering: (a) first  order, (b) second order and (c) third scattering order.}
  \label{fig:scatter_order}
\end{figure}

@@

To demonstrate recovery, we performed simulations, and tested effects of the following variations:
$\{1\}$ Multiple or only single scatter in the forward model;
$\{2\}$ Spatial distribution;
$\{3\}$ Density of viewpoints;
$\{4\}$ Aerosol characteristics;
$\{5\}$ Density scale.

In all cases, images were created as described
in \cref{sec:simul,sec:monte-carlo-simul}. These images served as input to the reconstruction algorithm. The optimization used an L-BFGS-B solver~\cite{BFGS} on a computer cluster. Each computer core was dedicated to rendering a modeled image. The optimization was initialized
by ${\bm n}=0$.  Satisfactory convergence occurred in several hundred
iterations. Depending on the resolution it took between minutes to a
couple of hours to complete, in total. The total estimation
error can be quantified by the aerosol mass that is over and under-estimated
in all voxels, relative to the total aerosol mass in the scene. Using the $\ell_1$ norm,
the total mass relative difference is
$(\|\hat{\bm n}\|_1 - \|{\bm n}^{\rm true} \|_1) / \| {\bm n}^{\rm true} \|_1$.
To also sense local errors, we use
 $\epsilon=\|
\hat{\bm n} - {\bm n}^{\rm true} \|_1 / \| {\bm n}^{\rm true} \|_1$.
%  In Figs.~\ref{fig:results}(a) and (b), respectively $\epsilon=0.14$ and
%$\epsilon=0.20$.  In Figs.~\ref{fig:results}(c) and (d), both cases have
%$\epsilon=0.65$.
These results are listed in Table~\ref{tbl:mass}. \\

\noindent{\bf Forward models and distributions}:
\Cref{fig:results} shows estimation results, corresponding to the
original distributions of \cref{fig:distributions}.
\begin{figure}[h]
  \centering
  \yoavcomment{\includegraphics{images/results}}
  \caption{\small Reconstructions of the scenes shown in
    \cref{fig:distributions}. Color represents aerosol density. The
    density units are $10^{6}~{\rm particles}/{\rm m}^3.$ In [a,b],
    single-scattering simulated images are used as input to the
    recovery.  In [c,d], MC simulated images are used as input for the
    recovery.  }
  \label{fig:results}
\end{figure}
Here, the atmosphere domain was defined over a
$20\times20 \times 40$ grid. Cameras were placed
$\sim 7$km apart on a $6\times6$ grid.
In Figs.~\ref{fig:results}(a) and (b), the single-scattering model was
used both in rendering of the raw images, and in the estimation algorithm. In
Figs.~\ref{fig:results}(c) and (d), the MC model (multiple scattering)
rendered the raw images, over which our recovery (Sec.~\ref{sec:inverse-problem}) was tested. Overall, the distributions, the density order of magnitude and
values are consistent.
Errors in Figs.~\ref{fig:results}(a) and (b) stem from random noise, which had
been induced in the raw images. Errors are larger when MC is used to
render the images, since high-order scattering is not accounted for in
the algorithm of \cref{sec:inverse-problem}.
\newcolumntype{g}{>{\columncolor{Gray}}r}
\begin{table}[htbp]
  \centering
  \begin{tabular}{cccccg}
    \toprule
    Scene  & Forward & Aerosol & Isotropic     & Absorbing & $\epsilon$ \\
           & model&   density & Phase function? & aerosol?   & \\
     \midrule
    Haze Front & Single scatter & Low & & & 0.18 \\
          & Monte Carlo & Low & & & 0.65 \\
    \cmidrule(r){1-5}
    Haze Blobs & Single scatter & Low & & & 0.14 \\
     & Monte Carlo & Low & & & 0.65 \\
     & Monte Carlo & Low & $\surd$  & & 0.49 \\
     & Monte Carlo & Low & & $\surd$ & 0.83 \\
     & Monte Carlo & High & & & 0.98 \\
    \bottomrule
  \end{tabular}
  \caption{Relative error measure in various simulations}
  \label{tbl:mass}
\end{table}
The overall mass is recovered, but for 6.5\% loss. \\

\noindent{\bf Density of viewpoints}: Keeping the domain and scene ({\em Haze blobs}) fixed, we
repeated the reconstruction many times, each time using a different subset of the above mentioned 36 simulated cameras. Each test randomly selected viewpoints, for various fixed values
of $N_{\rm views}< 36$. All rendered images used MC.
\Cref{fig:error_vs_cameras} plots $\epsilon$ as a function $N_{\rm views}$.
\begin{figure}[t]
  \centering
  \yoavcomment{\includegraphics{images/error_vs_cameras.pdf}}
  \caption{\small The relative error $\epsilon$ measure decreases with the number of
    cameras. Bars represent the standard deviation of $\epsilon$, over our random tests.}
  \label{fig:error_vs_cameras}
\end{figure}
Diminishing returns are obtained beyond $N_{\rm views}\approx 20$, which corresponds to $\approx 11$km separation between neighboring viewpoints.\\

\noindent{\bf Aerosol characteristics}: We used the {\em Haze blobs} scene, but varied the
aerosol characteristics. Compared to the aerosol described in \cref{sec:simul}, we tested two variations. In one variation, the phase function was isotropic, i.e.,
$[g_{\rm R},g_{\rm G},g_{\rm B}]=[0,0,0]$. In the other variation, the aerosol is partly absorbing: its single scattering albedo per color channel is
$[\varpi^{\rm aerosol}_{\rm R},\varpi^{\rm aerosol}_{\rm G},\varpi^{\rm aerosol}_{\rm B}]=
[0.69,0.74,0.78]$.  All rendered images used MC (See Figs.~\ref{fig:simulation_particles}(a)--(c)).
\begin{figure}[h]
  \centering
  \yoavcomment{\includegraphics{images/simulation_particles.pdf}}
  \caption{\small Images simulated by MC, from the same viewpoint, but different
  aerosol characteristics. (a) Haze blobs, characterized in Sec.~\ref{sec:simul}.
    high. (b) Partly absorbing aerosol. (c) Aerosol having an isotropic phase function. (d) High aerosol density.}
  \label{fig:simulation_particles}
\end{figure}
Recovery results using $N_{\rm views}=36$ are shown in Figs.~\ref{fig:reconstructions}(a,b) 
\begin{figure}[h]
  \centering
  \yoavcomment{\includegraphics{images/reconstructions.pdf}}
  \caption{\small  Recovered distributions when the aerosol is either party absorbing (a),
    isotropic (b) or highly dense (c).}
  \label{fig:reconstructions}
\end{figure}
and in Table~\ref{tbl:mass}.\\

\noindent{\bf Density scale}: We used the aerosol {\em Haze blobs} distribution
and characteristics as described in Sec.~\ref{sec:simul}, but increased the
aerosol density tenfold everywhere. This significantly increases the effects
of high order scattering, as shown in \cref{fig:cross_sections}(b) and
\cref{fig:simulation_particles}(d). Consequently, $\epsilon$ increases,
when using $N_{\rm views}=36$. Still, \cref{fig:reconstructions}(c) shows
that the spatial distribution outlines are recovered similarly to the other
tests.

@@

We describe a way to sense the 3D atmosphere. It can help remote
sensing depart from common 1D aerosol distribution models, in favor
of detailed 3D RT analysis.  The paper describes both a novel data
acquisition system concept for this recovery task (ground-based camera
network), and a dedicated algorithm for reconstructing aerosol
distributions. Using more prior knowledge on the nature of
distributions will improve the recovery.

The estimation algorithm
uses a single scattering model. This approximation is valid when the
optical depth is small. However, as MC simulations show, it is
important to generalize the recovery algorithm so that high-order scattering 
is included in it. Such generalization would be computationally complex: 
in our tests, the MC forward model was slower by 2-3 orders of magnitude than a single-scattering analytic model. Thus, inverting a 3D multiple-scattering model requires research into
numerical schemes for increasing the solution efficiency of both forward and inverse problems.  
The remote sensing community has devised several approaches to better approximate multiple-scattering problems, mainly in 1D scenes. These include successive orders of scattering, DISORT..... We may tap into some of these approaches.
Deploying an experimental system will be needed: camera specifications~\cite{Pust2011} should be
set to optimally recover a wide range of aerosol distributions.
\hl{The relation between the quantity and spatial distribution of
  cameras and the quality of reconstruction should also be determined.}
\fix{This requires}{These open questions require} a major theoretic
 and numerical research.

@@

BSF!!!!
